Stack Safety as a Security Property

### 2021-03-04 main.tex
Added sharing

Our stack-safety concepts depend critically on the notions of ``call'' and ``return,''
which are not fully explicit in machine code (the same instruction opcode may be
used in some places as part of a call or return sequence and in other places
for other purposes).  We define a call to be
any instruction declared as such in the compiler-provided call structure.
If a call occurs at $\PCname = a$ and $\rsp = s$, then the \emph{corresponding
  return} has occurred when the $\PCname$ points to the next instruction after $a$ and $\rsp = s$. (Our definition
of well-bracketed control flow relates this definition of return to the program's
declared set of return points.)
Our property definitions make no other detailed assumptions about call or return sequences, although
some of the enforcement mechanisms we consider do.

@@

Separate from the machine state, our formal properties will keep track of additional context
information representing the history of the run. The following definitions are parameterized
over an arbitrary set of context states, \(\CONTEXTS\), annotations \(\ANNS\),
an initial context state \(\context_0 \in \CONTEXTS\), and a context step function
\(\mach, \context, \ann \constep \context' \in \MACHS \times \CONTEXTS \times \ANNS \rightarrow \CONTEXTS\).

We lift contexts into an MP trace in the same way as we lifted policy states to get
a state type \(\mpcstatename \in \MPCS = \MACHS \times \POLS \times \CONTEXTS\).

\judgmenttwo{\(\mpstate{\mach}{\pol} \stepstopol[\obs] \mpstate{\mach'}{\pol'}\)}
            {\(\mach, \context \constep \context'\)}
            {\(\mpcstate{\mach}{\pol}{\context} \stepstocon[\obs]
               \mpcstate{\mach'}{\pol'}{\context'}\)}

@@

\subsection{Program Structure}

Our properties and policies assume that the program is annotated
by the compiler with information about its structure. At minimum,
we need to know which addresses contain instructions, which instructions
constitute calls and returns, and to which function an instruction
belongs.

This information is represented by a {\em code map}, a partial function from
addresses to pairs consisting of a function identifier and an {\em annotation}.
An annotation may be {\em call} or {\em return}.

\[\ann \in \ANNS ::= \callmap + \retmap\]
\[\codemap \in \CODEMAPS ::= \WORDS \rightharpoonup \FUNIDS \times \option(\ann)\]

@@

An initial context \(\context_0 = \domMap_0, 0\) where \(\domMap_0\) maps
the stack to \(\unsealed\) and all other addresses and all registers to \(\outside\).
Then at each step we update the context according to the context-step function,
\(\stepsto_C\), which we will define in terms of the operations
\(\push{\domMap}{n}{\components}\), in which \(\components \subseteq \COMPONENTS\),
and \(\pop{\domMap}{n}\), pronounced ``push'' and ``pop'' and defined as follows.

@@

\begin{center}
\begin{minipage}{.4\textwidth}
  \judgmenttwo{\(\mach \stepsto \mach'\)}
              {\(\codemap ~ (\mach ~ \PCname) = \funid, \mathbf{Some} ~ \callmap\)}
              {\(\mach,(\domMap,\depth) \stepstocon \mach',(\push{\domMap}{\depth}{\sealcon ~ \mach},\depth+1)\)}
\end{minipage}
\hspace*{0.1\textwidth}
\begin{minipage}{.4\textwidth}
  \judgmenttwo{\(\mach \stepsto \mach'\)}
              {\(\codemap ~ (\mach ~ \PCname) = \funid, \mathbf{Some} ~ \retmap\)}
              {\(\mach,(\domMap,\depth) \stepstocon \mach',(\pop{\domMap}{\depth},\depth-1)\)}
\end{minipage}
\end{center}

\begin{center}
  \judgment[Default]
           {\(\mach \stepsto \mach'\)}
           {\(\mach,\context \stepstocon \mach',\context\)}
\end{center}

@@

\[\ann \in \ANNS ::= \callmap, \retmap + (\sharemap ~ P \in \MACHS \times \COMPONENTS) +
(\passmap ~ P \in \MACHS \times \COMPONENTS)\]
\[\codemap \in \CODEMAPS ::= \WORDS \rightharpoonup \FUNIDS \times \powerset(\ann)\]

@@

@@

\begin{center}
  \judgmenttwo{\(\mach \stepsto \mach'\)}
              {\(\codemap ~ (\mach ~ \PCname) = \funid, \callmap \cup \passmap \cup \_\)}
              {\(\mach,(\domMap,\depth) \stepstocon (\push{\domMap}{\depth}{\sealcon ~ \mach},\depth+1)\)}
\hspace*{0.1\textwidth}

  \judgmenttwo{\(\mach \stepsto \mach'\)}
              {\(\codemap ~ (\mach ~ \PCname) = \funid, \mathbf{Some} ~ \retmap\)}
              {\(\mach,(\domMap,\depth) \stepstocon \mach',(\pop{\domMap}{\depth},\depth-1)\)}
\end{center}

\begin{center}
  \judgment[Default]
           {\(\mach \stepsto \mach'\)}
           {\(\mach,\context \stepstocon \mach',\context\)}
\end{center}

>>>>

Our stack-safety concepts depend critically on the notions of ``call'' and ``return,''
which are not fully explicit in machine code (the same instruction opcode may be
used in some places as part of a call or return sequence and in other places
for other purposes). We assume the compiler's cooperation in labeling instructions
with a variety of annotations, and in fact in identifying them as instructions.
A {\em call-map} is a partial map from addresses to pairs of a function identifier --
every instruction belongs to a function -- and a set of annotations. This gives
us flexibility to have instructions simultaneously play multiple roles.

@@

Separate from the machine state, our formal properties will keep track of additional context
information representing the history of the run. The following definitions are parameterized
over an arbitrary set of context states, \(\CONTEXTS\), annotations \(\ANNS\),
an initial context state \(\context_0 \in \CONTEXTS\), and a context step function
\(\mach, \context, \ann \in \powerset \ANNS \constep \context' \in \MACHS \times \CONTEXTS
\times \ANNS \rightarrow \CONTEXTS\). We also assume for any set of annotations that we have
code maps, with \(\FUNIDS\) a set of function identifiers:
\[\codemap \in \WORDS \rightharpoonup \FUNIDS \times \powerset(\ANNS)\]

We lift contexts into an MP trace in much the same way as we lifted policy states to get
a state type \(\mpcstatename \in \MPCS = \MACHS \times \POLS \times \CONTEXTS\).

\judgmentthree{\(\mpstate{\mach}{\pol} \stepstopol[\obs] \mpstate{\mach'}{\pol'}\)}
              {\(\codemap ~ (\mach ~ \PCname) = \funid, \ann\)}
              {\(\mach, \context, \ann \constep \context'\)}
              {\(\mpcstate{\mach}{\pol}{\context} \stepstocon[\obs]
                \mpcstate{\mach'}{\pol'}{\context'}\)}

@@

We begin with our only annotations being {\em call} or {\em return}, and
define our code map accordingly.

\[\ann \in \ANNS ::= \callmap + \retmap\]
\[\codemap \in \CODEMAPS ::= \WORDS \rightharpoonup \FUNIDS \times \powerset(\ANNS)\]

@@

An initial context \(\context_0 = \domMap_0, 0\) where \(\domMap_0\) maps
the stack to \(\unsealed\) and all other addresses and all registers to \(\outside\).
Then at each step we update the context according to the context-step function,
\(\stepsto_C\). We will define this in terms of the operations
\(\push{\domMap}{n}{\components}\), in which \(\components \subseteq \COMPONENTS\),
and \(\pop{\domMap}{n}\), pronounced ``push'' and ``pop'' and defined as follows.

@@

\begin{center}
\begin{minipage}{.4\textwidth}
  \judgment{}
           {\(\mach,(\domMap,\depth),\{\callmap\} \stepstocon
             \push{\domMap}{\depth}{\sealcon ~ \mach},\depth+1\)}
\end{minipage}
\hspace*{0.1\textwidth}
\begin{minipage}{.4\textwidth}
  \judgment{}
           {\(\mach,(\domMap,\depth),\{\retmap\} \stepstocon
             \pop{\domMap}{\depth},\depth-1\)}
\end{minipage}
\end{center}

\begin{center}
  \judgment[Default]
           {}
           {\(\_,\context,\_ \stepstocon \context\)}
\end{center}

@@

\[\ann \in \ANNS ::= \callmap + \retmap + \sharemap(P \subseteq \MACHS \times \COMPONENTS) +
\passmap(P \subseteq \MACHS \times \COMPONENTS)\]

@@

Finally we introduce a more direct update. If \(P\) is a relation between machine states and
components, we write \(\domMap[P ~ \mach \mapsto \stackDom]\) to say \(\domMap\) updated
so that each component \(\component\) where \(P ~ \mach ~ \component\) holds is mapped to
\(\stackDom\).

@@

\begin{center}
  \judgment{\(\mach,\context,\ann \stepstocon (\domMap,\depth)\)}
           {\(\mach,\context,(\{\callmap\} \cup \ann) \stepstocon
             (\push{\domMap}{\depth}{\sealcon ~ \mach},\depth+1)\)}
  \judgment{\(\mach,\context,\ann \stepstocon (\domMap,\depth)\)}
           {\(\mach,\context,(\{\retmap\} \cup \ann) \stepstocon
             \mach',(\pop{\domMap}{\depth},\depth-1)\)}
  \judgment{\(\mach,\context,\ann \stepstocon (\domMap,\depth)\)}
           {\(\mach,\context,(\{\passmap ~ P\} \cup \ann) \stepstocon
             \domMap[P ~ \mach \mapsto \passed{\depth}], \depth)\)}
  \judgment{\(\mach,\context,\ann \stepstocon (\domMap,\depth)\)}
           {\(\mach,\context,(\{\sharemap ~ P\} \cup \ann) \stepstocon
             \domMap[P ~ \mach \mapsto \shared{\depth}], \depth)\)}
  \judgment{}
           {\(\mach,\context,\_ \stepstocon \context\)}
\end{center}

\definition Let \(\MPCT\) be a reachable trace segment of the form
\(\mpcstatename ~ \mpcstatename'\), where \(\pi_\context(\mpcstatename) = (\domMap,\depth)\).
Let \(\components\) be the set of components \(\component\) such that either
\(\domMap ~ \component = \sealed{\_}\) or \(\domMap ~ \component = \passed{\depth'}\) for
some depth \(\depth' < \depth-1\). If any such \(\MPCT\) obeys trace integrity with respect
to \(\components\), then the system enjoys {\em ultra-eager confidentiality} in the sharing
model.

\definition Let \(\MPCT \langle f \rangle\) be a reachable trace segment where \(f\) holds
on contexts whose depth is less than or equal to some \(\depth\). Let \(\head(\MPCT) =
\mpcstate{\mach}{\pol}{(\domMap,\depth')}\) and \(\components\) be the set of components
\(\component\) such that either \(\domMap ~ \component = \sealed{\_}\) or
\(\domMap ~ \component = \passed{\depth''}\) for some depth \(\depth'' < \depth'-1\).
Then the system enjoys {\em eager integrity} in the sharing model if \(\MPCT\) obeys
trace integrity with respect to \(\components\).

\definition Let \(\MPCT \langle f \rangle\) be a reachable trace segment where \(f\) holds
on contexts whose depth is less than or equal to some \(\depth\). Let \(\head(\MPCT) =
\mpcstate{\mach}{\pol}{(\domMap,\depth')}\) and \(\components\) be the set of components
\(\component\) such that either \(\domMap ~ \component = \outside\) or
\(\domMap ~ \component = \shared{\_}\). Then the system enjoys {\em eager confidentiality}
in the sharing model if \(\MPCT\) obeys trace confidentiality with respect to \(\components\)
and \(f\).
